\chapter{Real-World Project Ideas}

Bridging the gap between theory and practice is essential. Here are five project ideas you can build using Hugging Face models, ranging from beginner to intermediate.

\section{1. Resume ATS Checker}
\textbf{The Problem:} Job applicants want to know if their resume matches a job description.
\\ \textbf{The Solution:} Use a \textbf{Zero-Shot Classification} model. These models can classify text into arbitrary categories without specific training examples.
\\ \textbf{Workflow:}
\begin{enumerate}
    \item \textbf{Input:} The full text of a candidate's resume (PDF to Text).
    \item \textbf{Labels:} Skills required (e.g., "Python", "Data Analysis", "Project Management").
    \item \textbf{Model:} \texttt{facebook/bart-large-mnli} (Robust and accurate).
    \item \textbf{Output:} A probability score for each skill. High scores mean the resume is a good match.
\end{enumerate}
\textbf{Key Library:} \texttt{transformers.pipeline("zero-shot-classification")}

\section{2. AI Chatbot Companion}
\textbf{The Problem:} Building a chatbot from scratch is hard and rule-based bots are boring.
\\ \textbf{The Solution:} Use a pre-trained conversational model trained on movie scripts or internet dialogue.
\\ \textbf{Workflow:}
\begin{enumerate}
    \item \textbf{Model:} \texttt{microsoft/DialoGPT-medium} or \texttt{facebook/blenderbot-400M-distill}.
    \item \textbf{Loop:} Take user input $\rightarrow$ Tokenize $\rightarrow$ Generate Response $\rightarrow$ Decode.
    \item \textbf{Challenge:} Keep a "history" list of past inputs so the bot remembers the conversation context. Without history, the bot has short-term memory loss.
\end{enumerate}

\section{3. Multilingual Review Translator}
\textbf{The Problem:} An app has reviews in Spanish, French, and Hindi, but the analytics team only speaks English.
\\ \textbf{The Solution:} A universal translation pipeline.
\\ \textbf{Workflow:}
\begin{enumerate}
    \item \textbf{Language ID:} Use \texttt{papluca/xlm-roberta-base-language-detection} to detect the language.
    \item \textbf{Translate:} Use \texttt{facebook/m2m100\_418M} (Many-to-Many) to translate any language to English.
    \item \textbf{Analyze:} improved sentiment analysis on the translated text.
\end{enumerate}

\section{4. "Too Long; Didn't Read" News Summarizer}
\textbf{The Problem:} News articles are too long and full of fluff.
\\ \textbf{The Solution:} An abstractive summarizer that rewrites the core meaning.
\\ \textbf{Workflow:}
\begin{enumerate}
    \item \textbf{Scraper:} Use \texttt{BeautifulSoup} to get text from a URL.
    \item \textbf{Model:} \texttt{sshleifer/distilbart-cnn-12-6} (Fast and light).
    \item \textbf{Settings:} Set \texttt{max\_length=60} to force a conciseness.
\end{enumerate}

\section{5. Intelligent Flashcard Generator}
\textbf{The Problem:} Creating study notes takes time.
\\ \textbf{The Solution:} Convert textbook photos into Q\&A pairs using Vision + Language models.
\\ \textbf{Workflow:}
\begin{enumerate}
    \item \textbf{OCR:} Use \texttt{microsoft/trocr-base-handwritten} to read text from an image.
    \item \textbf{Question Generation:} Use a T5 model fine-tuned on "answer2question" tasks (like \texttt{mrm8488/t5-base-finetuned-question-generation-ap}).
    \item \textbf{Output:} A JSON file of Question-Answer pairs ready for Anki.
\end{enumerate}

\begin{examplebox}{Project Flow Visualization}
    \begin{center}
    Resume Text \\ $\downarrow$ \\ \textbf{Model Layer} (BART-MNLI) \\ $\downarrow$ \\ \textit{"Python: 99\%, Java: 12\%"}
    \end{center}
\end{examplebox}
